---
title: "Nobel winners"
author: ""
output: github_document
---

```{r load-packages, message=FALSE}
library(tidyverse)
```

1. Let's first load the nobels.csv file from the data-raw/ folder:

```{r label-me1, eval=FALSE}
nobel <- read_csv("data-raw/nobel.csv")
```

2. Then let's split the data into two new data frames called `nobel_stem` (which filters for the STEM fields: Physics, Medicine, Chemistry, and Economics) and `nobel_nonstem` that includes the remaining fields. 
```{r label-me2, eval=FALSE}
# stem laureates
nobel_stem <- nobel %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))

# non-steam laureates
nobel_nostem <- nobel %>%
  filter(!category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

3. And finally write the resulting data frames to .csv files called nobel-stem.csv and nobel-nostem.csv to the data/ folder.
```{r label-me3}
write_csv(nobel_stem, "data-raw/nobel_stem.csv")
write_csv(nobel_nostem, "data-raw/nobel_nostem.csv")
```
