---
title: "Bechdel analysis"
author: "___"
output: 
  github_document: 
    fig_height: 5
    fig_width: 8
---

In this mini analysis we work with the data used in the FiveThirtyEight story titled ["The Dollar-And-Cents Case Against Hollywood’s Exclusion of Women"](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/). According to the FiveThirtyEight article, "if a movie can satisfy three criteria — there are at least two named women in the picture, they have a conversation with each other at some point, and that conversation isn’t about a male character — then it passes “The Rule,” whereby female characters are allocated a bare minimum of depth.

Your task is to fill in the blanks denoted by `___`. If the `___` appears in a code chunk, remove the `#` at the beginning of the line as well. The `#` creates a comment, which is not evaluated by R. We want the code to be evaluated once you fix it.

## Data and packages

We start with loading the packages we'll use.

```{r load-packages, message=FALSE}
library(fivethirtyeight)
library(tidyverse)
```

The dataset contains information on `r nrow(bechdel)` movies released between `r min(bechdel$year)` and `r max(bechdel$year)`. However we'll focus our analysis on movies released between 1990 and 2013.

```{r}
bechdel90_13 <- bechdel %>% 
  filter(between(year, 1990, 2013))
```

1. There are `___` movies between 1990 and 2013. Hint: each row in the `bechdel90_13` table we just made represents a different movie. Type `bechdel90_13` into the console to see the table dimensions.

The financial variables we'll focus on are the following:

- `budget_2013`: Budget in 2013 inflation adjusted dollars
- `domgross_2013`: Domestic gross (US) in 2013 inflation adjusted dollars
- `intgross_2013`: Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

And we'll also use the `binary` and `clean_test` variables for **grouping**.

## Analysis

Let's take a look at how median budget and gross vary by whether the movie passed the Bechdel test, which is stored in the `binary` variable.

```{r message = FALSE}
bechdel90_13 %>%
  group_by(binary) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

Next, let's take a look at how median budget and gross vary by a more detailed indicator of the Bechdel test result. This information is stored in the `clean_test` variable, which takes on the following values:

- `ok` = passes test
- `dubious`
- `men` = women only talk about men
- `notalk` = women don't talk to each other
- `nowomen` = fewer than two women

2. Remove the `#` and replace the `___` with the correct appropriate grouping variable.
```{r message = FALSE}
bechdel90_13 %>%
  #group_by(___) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

In order to evaluate how return on investment varies among movies that pass and fail the Bechdel test, we'll  first create a new variable called `roi` as the ratio of the gross to budget.

```{r}
bechdel90_13 <- bechdel90_13 %>%
  mutate(roi = (intgross_2013 + domgross_2013) / budget_2013)
```

Let's see which movies have the highest return on investment.

```{r}
bechdel90_13 %>%
  arrange(desc(roi)) %>% 
  select(title, roi, year)
```

3. Below is a visualization of the return on investment by test result. Enter an appropriate y-axis label in the code below. 

```{r warning = FALSE}
ggplot(data = bechdel90_13, 
       mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       x = "Detailed Bechdel result",
       y = "___",
       color = "Binary Bechdel result")
```

However it's difficult to see the distributions due to a few extreme observations. The following movies have *very* high returns on investment.

```{r}
bechdel90_13 %>%
  filter(roi > 400) %>%
  select(title, budget_2013, domgross_2013, year)
```

4. Zooming in on the movies with `roi < ___` provides a better view of how the medians across the categories compare:

5. Enter an appropriate y-axis label in the code below. 
```{r warning = FALSE}
ggplot(data = bechdel90_13, mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       subtitle = "___", # Something about zooming in to a certain level
       x = "Detailed Bechdel result",
       y = "Return on investment",
       color = "Binary Bechdel result") +
  coord_cartesian(ylim = c(0, 20))
```

6. What does this plot suggest about the relationship between Bechdel test results and movie return on investment?
